---
title: "Black sea bass ecosystem considerations"
output: html_document
date: "`r Sys.Date()`"
author: "Abigail Tyrell"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
`%>%` <- magrittr::`%>%`
```

## Data

### Independent variables

#### Bottom temperature
```{r}
dat <- read.csv(here::here("data/bsb_bt_temp_nmab_1959-2022.csv"))
sdat <- read.csv(here::here("data/bsb_bt_temp_smab_1959-2022.csv"))

dat2 <- dat %>%
  dplyr::bind_rows(sdat) %>%
  dplyr::rename(val = mean,
                name = Region) %>%
  dplyr::group_by(name) %>%
  dplyr::mutate(mean = mean(val),
                sd = sd(val)) %>%
  dplyr::ungroup()
```

#### Shelf water volume
```{r}
n_swv <- readxl::read_excel(here::here("data-raw/ShelfWaterVolume_BSB_Update.xlsx"),
                            sheet = "N. MAB") %>%
  janitor::clean_names() %>%
  dplyr::select(year,
                val = sh_w_vol) %>%
  dplyr::mutate(region = "North")

s_swv <- readxl::read_excel(here::here("data-raw/ShelfWaterVolume_BSB_Update.xlsx"),
                            sheet = "S. MAB") %>%
  janitor::clean_names() %>%
  dplyr::select(year,
                val = sh_w_vol) %>%
  dplyr::mutate(region = "South")

swv <- rbind(n_swv,
             s_swv) %>%
  dplyr::mutate(whole_year = trunc(year),
                dec_year = year - whole_year) %>%
  dplyr::filter(dec_year < 0.25) %>% # winter only
  dplyr::select(-year, -dec_year) %>%
  dplyr::rename(year = whole_year,
                name = region) %>%
  dplyr::group_by(year, name) %>%
  dplyr::summarise(new_val = mean(val)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(name) %>%
  dplyr::mutate(mean = mean (new_val),
                sd = sd(new_val)) %>%
  dplyr::rename(val = new_val)
```

### Dependent variables

#### Recruitment
```{r}
n_rec <- read.table(here::here("data-raw/NORTH.MT.2021.FINAL.STD.txt"),
                    header = TRUE) %>%
  dplyr::filter(name == "log_recruit_devs") %>%
  dplyr::mutate(year = 1989:2019) %>%
  dplyr::rename(val = value)

s_rec <- read.table(here::here("data-raw/SOUTH.MT.2021.FINAL.STD.txt"),
                    header = TRUE) %>%
  dplyr::filter(name == "log_recruit_devs") %>%
  dplyr::mutate(year = 1989:2019) %>%
  dplyr::rename(val = value)

rec <- rbind(n_rec %>%
               dplyr::mutate(region = "North"),
             s_rec %>%
               dplyr::mutate(region = "South")) %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(mean = mean(val),
                sd = sd(val))
```

#### Effective area
```{r}
load(here::here("data/from_Alex/parameter_estimates.RData"))
Sdreport_spring <-parameter_estimates$SD

nrow_sd <- length(Sdreport_spring$value)
year<-1989:2021
n_yea <- length(unique(year))
n_surv <- 11
n_stock<-2

df_sdreport_spring <-
  data.frame(variable = c(names(Sdreport_spring$value)),
             value = c(Sdreport_spring$value),
             sd = c(Sdreport_spring$sd),
             Season = c(rep("Spring", times = nrow_sd)))

df_effArea <-
  df_sdreport_spring %>%
  dplyr::filter(variable %in% c("effective_area_ctl"))%>%
  dplyr::mutate(Region = rep(c("North", "South"),
                           times = n_yea*n_surv))%>%
  dplyr::mutate(year   = rep(c(1989:2021),
                             each = 2,
                             times = 11),
                survey   = rep(c(1:11),
                               each = 66,
                               times = 1))
# 
# df_effArea <- data.frame(variable = names(parameter_estimates$SD$value),
#                          value = parameter_estimates$SD$value,
#                          sd = parameter_estimates$SD$sd) %>%
#   dplyr::filter(variable == "effective_area_ctl") %>%
#   dplyr::mutate(size = rep(c("North", "South"),
#                            times = length(1989:2021)*11),
#                 year = rep(c(1989:2021),
#                            each = 2,
#                            times = 11),
#                 survey = rep(c(1:11),
#                              each = 66)) %>%
#   dplyr::filter(survey == 1)

df_effArea2 <- df_effArea %>%
  dplyr::filter(survey == 1) %>%
  dplyr::rename(val = value,
                region = Region) %>%
  dplyr::select(val, region, year) %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(mean = mean(val),
                sd = sd(val),
                name = "eff_area")
```

#### Center of gravity
```{r}
df_cog_spring <-
  df_sdreport_spring %>%
  dplyr::filter(variable %in% c("mean_Z_ctm"))%>%
  dplyr::mutate(variable = rep(c("Eastings", "Northings"),
                               each = n_yea*n_stock),
                Region = rep(c("North", "South"),
                            each = 1,
                            times = n_yea*n_stock),
                year   = rep(c(1989:2021),
                             each = 2,
                             times = 2)) %>%
  dplyr::rename(region = Region,
                val = value,
                annual_sd = sd) %>%
  dplyr::group_by(region, variable) %>%
  dplyr::mutate(mean = mean(val),
                sd = sd(val),
                name = "cog")
```

## Functions
```{r}
time_series_plot <- function(data, ylab) {
  plt <- data %>%
    ggplot2::ggplot(
      ggplot2::aes(
        x = .data$year,
        y = .data$val,
        group = name
      )
    ) +
    ggplot2::geom_hline(ggplot2::aes(
      yintercept = .data$mean + .data$sd,
      group = .data$name
    ),
    color = "darkgreen",
    linetype = "solid"
    ) +
    ggplot2::geom_hline(ggplot2::aes(
      yintercept = .data$mean - .data$sd,
      group = .data$name
    ),
    color = "darkgreen",
    linetype = "solid"
    ) +
    ggplot2::geom_hline(ggplot2::aes(
      yintercept = .data$mean,
      group = .data$name
    ),
    color = "darkgreen",
    linetype = "dotted"
    ) +
    ggplot2::geom_point() +
    ggplot2::geom_line(data = data %>%
                         tidyr::drop_na(.data$val)
    ) +
    ggplot2::ylab(ylab) +
    ggplot2::scale_y_continuous(labels = scales::comma) +
    ggplot2::theme_bw(base_size = 16) +
    ggplot2::theme(strip.text = ggplot2::element_text(size = 16),
                   axis.title.x = ggplot2::element_blank())

  return(plt)

}

test_cor <- function(data, y1, y2 = NULL, x, alt) {
  for(i in c("North", "South")) {
    message(paste(i, y1))
    res <- cor.test(x = data %>%
                      dplyr::filter(.data$region == i) %>%
                      dplyr::pull(x),
                    y = data %>%
                      dplyr::filter(.data$region == i) %>%
                      dplyr::pull(y1),
                    alternative = alt)
    
    out <- tibble::tibble("Region" = i,
                          "Independent variable" = x,
                          "Dependent variable" = y1,
                          "Correlation coefficient" = res$estimate,
                          "P value" = res$p.value) %>%
      t()
    out %>%
      knitr::kable() %>%
      print()
    
    if(!is.null(y2)) {
          message(paste(i, y2))
    res <- cor.test(x = data %>%
                      dplyr::filter(.data$region == i) %>%
                      dplyr::pull(x),
                    y = data %>%
                      dplyr::filter(.data$region == i) %>%
                      dplyr::pull(y2),
                    alternative = alt)
    
        out <- tibble::tibble("Region" = i,
                          "Independent variable" = x,
                          "Dependent variable" = y2,
                          "Correlation coefficient" = res$estimate,
                          "P value" = res$p.value) %>%
          t()
    out %>%
      knitr::kable() %>%
      print()
    }
  }
}

compare_norm <- function(data, 
                         ... # passed to ggplot aesthetics
                         ) {
  fig <- data %>%
  ggplot2::ggplot(ggplot2::aes_string(...)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::facet_wrap(~region,
                      nrow = 2) +
  ggplot2::xlim(c(1989, NA)) +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::theme(legend.position = "bottom",
                 legend.title = ggplot2::element_blank(),
                 strip.text = ggplot2::element_text(size = 16),
                 axis.title.x = ggplot2::element_blank()) +
  ggplot2::ylab("Normalized value")
  
  return(fig)
}

cor_plt <- function(data,
                    ..., # passed to aes
                    xlab,
                    ylab
                    ) {
  fig <- data %>%
  ggplot2::ggplot(ggplot2::aes_string(...)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm") +
  viridis::scale_color_viridis() +
    ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::facet_wrap(~region,
                      scales = "free",
                      ncol = 1) +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::theme(legend.position = "right",
                 legend.direction = "vertical") +
  ggplot2::ylab(ylab) +
  ggplot2::xlab(xlab)
  
  return(fig)
}
```

## Plots & statistics

### Time series

#### Bottom temperature
```{r}
time_series_plot(data = dat2, ylab = "Winter bottom temperature") +
  ggplot2::facet_wrap(~name,
                      nrow = 2)
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_temperature.png")),
                           width = 5, height = 6)
```

#### Shelf water volume
```{r}
swv %>%
  time_series_plot(ylab = "Winter shelf water volume") +
  ggplot2::facet_wrap(~name,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_swv.png")),
                width = 5, height = 6)
```

#### Recruitment
```{r}
rec %>%
        time_series_plot(ylab = "Log Recruitment Deviations (2021 model)") +
  ggplot2::facet_wrap(~region,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_recruitment.png")),
                 width = 5, height = 6)
```

#### Effective area
```{r}
df_effArea2 %>%
  time_series_plot(ylab = "Effective area (km2)") +
  ggplot2::facet_wrap(~region,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_effective_area.png")),
                width = 5, height = 6)

```

#### Center of gravity

##### Mean COG
```{r}
df_cog_spring %>%
  dplyr::filter(variable == "Northings") %>%
  time_series_plot(ylab = "Mean northings (km)") +
  ggplot2::facet_wrap(~region,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_northings.png")),
                 width = 5, height = 6)
```

##### Lower CI COG
```{r}
df_cog_spring %>%
  dplyr::filter(variable == "Northings") %>%
  dplyr::mutate(val = val - 1.96*annual_sd) %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(mean = mean(val),
                sd = sd(val)) %>%
  time_series_plot(ylab = "Lower CI northings (km)") +
  ggplot2::facet_wrap(~region,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_northings_lower.png")),
                 width = 5, height = 6)
```

##### Upper CI COG
```{r}
df_cog_spring %>%
  dplyr::filter(variable == "Northings") %>%
  dplyr::mutate(val = val + 1.96*annual_sd) %>%
    dplyr::group_by(region) %>%
  dplyr::mutate(mean = mean(val),
                sd = sd(val)) %>%
  time_series_plot(ylab = "Upper CI northings (km)") +
  ggplot2::facet_wrap(~region,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_northings_upper.png")),
                 width = 5, height = 6)
```

##### North vs South COG
```{r}
df_cog_spring %>%
  dplyr::filter(variable == "Northings") %>%
  ggplot2::ggplot(ggplot2::aes(x = year,
                               y = val,
                               fill = region,
                               color = region)) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = val - 1.96*annual_sd,
                                    ymax = val + 1.96*annual_sd),
                       alpha = 0.2,
                       color = NA) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::ylab("Northings (km)")

ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_northings_ci.png")),
                 width = 8, height = 6)
```

##### Eastings (mean COG)
```{r}
df_cog_spring %>%
  dplyr::filter(variable == "Eastings") %>%
  time_series_plot(ylab = "Eastings (km)") +
  ggplot2::facet_wrap(~region,
                      nrow = 2,
                      scales = "free_y")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_eastings.png")),
                 width = 5, height = 6)
```

### Correlations

#### Bottom temperature x recruitment
```{r}
compare <- rbind(rec,
                 dat2 %>%
                   dplyr::rename(region = name) %>%
                   dplyr::mutate(name = "bottom temperature")) %>%
  dplyr::filter(year > 1988) %>%
                 dplyr::mutate(norm_value = (val - mean)/sd)
```

```{r}
cor_dat <- compare %>%
  dplyr::select(name, val, region, year) %>%
  dplyr::mutate(name = ifelse(name == "bottom temperature", "bottom_temp", name)) %>%
  tidyr::pivot_wider(names_from = name,
                     values_from = val)

cor_dat %>%
    cor_plt(x = "bottom_temp",
          y = "log_recruit_devs",
          color = "year",
          ylab = "Log recritment deviation",
          xlab = "Winter bottom temperature")
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "temp_x_rec2.png")),
                width = 8, height = 6)
```

```{r}
compare %>%
  compare_norm(x = "year",
               y = "norm_value",
               color = "name")
  
ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_temp_x_rec.png")),
                width = 5, height = 6)
```

```{r, results = "asis"}
test_cor(data = cor_dat,
         x = "bottom_temp",
         y1 = "log_recruit_devs",
         alt = "two.sided")
```

#### Shelf water volume x effective area
```{r}
dat <- dplyr::full_join(df_effArea2 %>%
                          dplyr::select(-c(mean, sd, name)) %>%
                          dplyr::rename(eff_area = val),
                        swv %>%
                          dplyr::select(-c(mean, sd)) %>%
                          dplyr::rename(swv = val,
                                        region = name)) %>%
  tidyr::drop_na()

dat %>%
  cor_plt(x = "swv",
          y = "eff_area",
          color = "year",
          ylab = "Spring effective area occupied (km2)",
          xlab = "Winter shelf water volume (km3)")

ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_swv_x_eff_area.png")),
                width = 8, height = 6)
```

```{r}
compare <- rbind(df_effArea2,
                 swv %>%
                   dplyr::rename(region = name) %>%
                   dplyr::mutate(name = "swv"))%>%
                 dplyr::mutate(norm_value = (val - mean)/sd)
compare %>%
  compare_norm(x = "year",
               y = "norm_value",
               color = "name")

ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_swv_x_eff_area2.png")),
                width = 8, height = 6)
```

```{r, results = "asis"}
test_cor(data = dat, 
         y1 = "eff_area", 
         x = "swv", 
         alt = "two.sided")
```

#### Shelf water volume x center of gravity
```{r}
dat <- dplyr::full_join(df_cog_spring %>%
                          dplyr::filter(variable == "Northings") %>%
                          dplyr::select(region, year, val) %>%
                          dplyr::rename(northings = val),
                        swv %>%
                          dplyr::select(-c(mean, sd)) %>%
                          dplyr::rename(swv = val,
                                        region = name)) %>%
  tidyr::drop_na()

dat %>%
  cor_plt(x = "swv",
          y = "northings",
          color = "year",
          ylab = "Spring northings (km)",
          xlab = "Winter shelf water volume (km3)")

ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_swv_x_northings.png")),
                width = 8, height = 6)
```

```{r}
compare <- rbind(df_cog_spring %>%
                          dplyr::filter(variable == "Northings") %>%
                   dplyr::mutate(name = "northings"),
                 swv %>%
                   dplyr::rename(region = name) %>%
                   dplyr::mutate(name = "swv"))%>%
                 dplyr::mutate(norm_value = (val - mean)/sd)
compare %>%
  compare_norm(x = "year",
               y = "norm_value",
               color = "name")

ggplot2::ggsave(here::here("images", paste0(Sys.Date(), "_swv_x_northings2.png")),
                width = 8, height = 6)
```

```{r, results = "asis"}
test_cor(data = dat, 
         y1 = "northings", 
         x = "swv", 
         alt = "two.sided")
```
