---
title: "BSB survival"
date: "2024-07-31"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width=12)
```


```{r load_dat, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(lubridate)

# remotes::install_github("NOAA-EDAB/stocksmart")
# install.packages("dsem")
# library(dsem) ## 1.2.1
# library(phylopath)
library(stocksmart)
library(mgcv)
library(gratia)
source(here::here("scripts/thresholds_gam.R"))



bt_dat <- read.csv(here::here("data/bt_update_2024-05-06.csv")) %>% 
  filter(Region %in% c("North", "South")) %>% 
  mutate(date = as.Date(paste0(year, "-01-01")),
         upper = mean + se,
         lower = mean - se)

recruit_dat_s <- read.csv(here::here("data/1.south.Rect.yr.csv")) %>% 
  mutate(date = as.Date(paste0(Year, "-01-01")),
         Region = "South")

recruit_dat_n <- read.csv(here::here("data/1.north.Rect.yr.csv")) %>% 
  mutate(date = as.Date(paste0(Year, "-01-01")),
         Region = "North")
recruit_dat <- bind_rows(recruit_dat_s, recruit_dat_n)

# ssb_dat_n <- read.csv("https://raw.githubusercontent.com/emilyliljestrand/BSB.2024.MT/main/Modeling/Run3/fit/outputs/1.north.SSB.yr.csv") %>% 
#   mutate(date = as.Date(paste0(Year, "-01-01")),
#          Region = "North")
# saveRDS(ssb_dat, file = "data/ssb_dat.rds")
# ssb_dat_s <- read.csv("https://raw.githubusercontent.com/emilyliljestrand/BSB.2024.MT/main/Modeling/Run3/fit/outputs/1.south.SSB.yr.csv?token=GHSAT0AAAAAACVRCJTCC24RZMMFAMPB2Q52ZVKNWRA") %>% 
#   mutate(date = as.Date(paste0(Year, "-01-01")),
#          Region = "South")
# ssb_dat <- bind_rows(ssb_dat_n, ssb_dat_s)
ssb_dat <- readRDS(here::here("data/ssb_dat.rds"))
```


## R Markdown


```{r bt_plot}

bt_plot <- ggplot(bt_dat,aes(x = date, y = mean, ymin = lower, ymax = upper, color = Region, fill = Region), show.legend = FALSE) +
  geom_line(alpha = 0.3)+
  geom_point()+
  theme_minimal() +
  scale_x_date(date_breaks = "5 years",
               date_labels = "%Y") +
  labs(title = "Winter bottom temperature (°C)",
       subtitle = "GLORYS reanalysis",
       x = "",
       y = "",
       fill = "", 
       color = "") +
  theme(legend.position = "bottom")
bt_plot
```


```{r recruit_plot}

recruit_plot <- ggplot(recruit_dat,
                  aes(x = date, y = est, color = Region)) +
  geom_path() +
  geom_point() +
  theme_minimal() +
  scale_x_date(date_breaks = "5 years",
               date_labels = "%Y") +
  labs(title = "Black Sea Bass recruits",
       subtitle = "WHAM output",
       x = "",
       y = "",
       color = "") +
  theme(legend.position = "bottom")
recruit_plot
```


```{r ssb_plot}

ssb_plot <- ggplot(ssb_dat,
                  aes(x = date, y = est, color = Region)) +
  geom_path() +
  geom_point() +
  theme_minimal() +
  scale_x_date(date_breaks = "5 years",
               date_labels = "%Y") +
  labs(title = "Black Sea Bass SSB",
       subtitle = "WHAM output",
       x = "",
       y = "",
       color = "") +
  theme(legend.position = "bottom")
ssb_plot
```



```{r recruit_bt}
recruit_bt_dat <- bt_dat %>% 
  select(date, Region, bt = mean) %>% 
  left_join(recruit_dat %>% 
              select(date, Region, recruits = est), by = join_by(date, Region)) %>% 
  mutate(year = lubridate::year(date))

recruit_bt_plot <- ggplot(recruit_bt_dat, aes(x = bt, y = recruits, color = Region, label = year)) +
  # geom_path() + 
  geom_point(show.legend = FALSE) +
  geom_text(show.legend = FALSE, vjust = "bottom") +
  theme_minimal() +
  facet_wrap(~Region, ncol = 2) +
  labs(title = "Black Sea Bass recruitment and bottom temperature",
       x = "Bottom temperature (°C)",
       y = "Black Sea Bass recruits",
       color = "") +
  # theme(legend.position = "bottom") +
  NULL

recruit_bt_plot

```



```{r ssb_recruit_bt}
ssb_recruit_bt_dat <- bt_dat %>% 
  select(date, Region, bt = mean) %>% 
  left_join(recruit_dat %>% 
              select(date, Region, recruits = est), by = join_by(date, Region)) %>% 
  mutate(year = lubridate::year(date)) %>% 
  left_join(ssb_dat %>% 
              select(date, Region, ssb = est), by = join_by(date, Region)) 

ssb_recruit_bt_plot <- ggplot(ssb_recruit_bt_dat, aes(x = ssb, y = recruits, size = bt, color = bt, label = year)) +
  # geom_path() + 
  geom_point(aes(alpha = bt), show.legend = TRUE) +
  geom_text(show.legend = FALSE, vjust = "bottom", color = "black") +
  theme_minimal() +
  facet_wrap(~Region, ncol = 2) +
  guides(size = "none", alpha = "none") + 
  scale_color_viridis_c(option = "inferno", direction = -1) +
  labs(title = "Black Sea Bass SSB, recruitment, and bottom temperature",
       x = "Black Sea Bass SSB",
       y = "Black Sea Bass recruits",
       size = "",
       color = "bottom temperature") +
  theme(legend.position = "bottom") +
  NULL

ssb_recruit_bt_plot

```




## GAM Thresholds


### recruitment
1. recruits = s(year)
2. recruits = s(year) if BT < th


```{r gam_model}
##North
n_dat <- ssb_recruit_bt_dat %>%
  filter(Region == "North") %>%
  mutate(rec_spawner = recruits/ssb) %>%
  as.data.frame()
# n_1 <- mgcv::gam(log(recruits) ~ s(bt) + s(ssb) + s(year), data = n_dat)
# gratia::draw(n_1)
# gratia::appraise(n_1)
# summary(n_1)
# acf(resid(n_1))

# source("https://raw.githubusercontent.com/duncanobrien/ews-assessments/main/Code/threshold_gam.R")
# https://github.com/duncanobrien/ews-assessments/blob/main/Code/threshold_gam.R
  # data, #data.frame
  #           threshold.name, # colname of threshold variable to test within data
  #           nthd, #number of steps between lower & upper bound
  #           a = 0.2, #probability for lower bound quantile, 0:1
  #           b = 0.8, #probability for upper bound quantile, 0:1

# n_2 <- threshold.gam(log(recruits) ~ s(ssb) + s(year), nthd = 10, threshold.name = "bt", data = n_dat)
# summary(n_2)
# data = n_dat
cont_formula_ts <- as.formula("recruits ~ s(year, bs='tp', k=3)")
thresh_formula_ts <- as.formula("recruits ~ 1")

dens_gam_ts <- compare_gam(data = n_dat, 
                        cont_formula = formula(cont_formula_ts, method = "REML"),
                        thresh_formula =  formula(thresh_formula_ts, method = "REML"),
                        thresh.var = "year",
                        thresh.range = c(0.15,0.85), by = 1, k=3) #quartile range for threshold within 15-85%. Prevents bias towards thresholds at ends of time series


best_gam_ts <- predict_best_gam(object = dens_gam_ts) %>%
  dplyr::rename("metric.val" = recruits) %>% 
  dplyr::mutate(threshold = tidyr::replace_na(as.character(threshold),"pre"),
                                                    start.date = dplyr::first(thresh.var),
                                                    last.date = dplyr::last(thresh.var)) %>% 
  pivot_longer(year, names_to = "var_metric", values_to = "var.value")


cont_formula_ss <- as.formula("recruits ~ s(bt, bs='tp', k=3)")
thresh_formula_ss <- as.formula("recruits ~ 1")
dens_gam_ss <- compare_gam(data = n_dat,
                        cont_formula = formula(cont_formula_ss, method = "REML"),
                        thresh_formula =  formula(thresh_formula_ss, method = "REML"),
                        thresh.var = "year", expl.var = "bt",
                        thresh.range = c(0.2,0.75), by = 1, k=3)

best_gam_ss <- predict_best_gam(object = dens_gam_ss) %>%
  dplyr::rename("metric.val" = recruits) %>% 
  dplyr::mutate(threshold = tidyr::replace_na(as.character(threshold),"pre"),
                start.date = dplyr::first(thresh.var),
                last.date = dplyr::last(thresh.var)) %>% 
    pivot_longer(bt, names_to = "var_metric", values_to = "var.value")


best_gam <- bind_rows(best_gam_ts, best_gam_ss) %>% 
   mutate(var_metric = factor(ifelse(var_metric == "bt","Bottom Temperature","Year"),levels = c("Year","Bottom Temperature"), labels = c("Time series","State space")))

ts_plot <- ggplot(data = subset(best_gam, var_metric == "Time series"), aes(x=var.value,y=metric.val)) + 
  geom_point(aes(x=thresh.var, y = metric.val))+
  geom_path(aes(x=thresh.var, y = metric.val)) +  
  xlab("Year") + ylab("Scaled metric score") +
  facet_wrap(~ var_metric, scales = "free_x") +
  geom_line(data = filter(best_gam, threshold=="pre" & var_metric == "Time series"),aes(x=var.value,y=fit), col="blue",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="pre"& var_metric == "Time series"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#A1B4FE", col="#A1B4FE",alpha = 0.2) +
  geom_line(data = filter(best_gam, threshold=="post"& var_metric == "Time series"),aes(x=var.value,y=fit), col="red",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="post"& var_metric == "Time series"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#FFA6B9", col="#FFA6B9",alpha = 0.2)+
  geom_point(data = filter(best_gam, thresh.var == start.date & var_metric == "Time series"),aes(x=var.value, y = metric.val,col="Start date"))+
  geom_point(data = filter(best_gam, thresh.var == last.date & var_metric == "Time series"),aes(x=var.value, y = metric.val,col="End date"))+
  geom_point(data = filter(best_gam,transition == "trans" & var_metric == "Time series"),aes(x=var.value, y = metric.val,col="Transition\ndates"))+
  ggrepel::geom_text_repel(data = filter(best_gam, transition == "trans" & var_metric == "Time series"), aes(x=var.value, y = metric.val,label=thresh.var),force =1.5,nudge_x=5,nudge_y=-0.75,segment.linetype=2,min.segment.length = 0.1)+
  scale_colour_manual(values = c("blue","red","#FFE823"),breaks = c("Start date", "End date", "Transition\ndates"), name = NULL, guide = guide_legend(override.aes = list(size = 5))) +
  theme_classic() +
  scale_x_continuous(breaks = seq(1970,2010,20)) +
  theme(legend.text=element_text(size=14),
        strip.text.y = element_blank(),
        strip.text.x = element_text(size=14),
        axis.title = element_text(size=14),
        legend.position = "top",
        panel.border = element_rect(linewidth = 1,fill = "transparent"))
                                        

ss_plot <- ggplot(data = subset(best_gam,var_metric == "State space"), aes(x=var.value,y=metric.val)) + 
  geom_point(aes(x=var.value, y = metric.val))+
  geom_path(aes(x=var.value, y = metric.val)) +  
  xlab("Environment score") + ylab("Scaled metric score") +
  facet_wrap(~ var_metric, scales = "free_x") +
  geom_line(data = filter(best_gam, threshold=="pre" & var_metric == "State space"),aes(x=var.value,y=fit), col="blue",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="pre"& var_metric == "State space"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#A1B4FE", col="#A1B4FE",alpha = 0.2)+
  geom_line(data = filter(best_gam, threshold=="post"& var_metric == "State space"),aes(x=var.value,y=fit), col="red",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="post"& var_metric == "State space"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#FFA6B9", col="#FFA6B9",alpha = 0.2)+
  geom_point(data = filter(best_gam, thresh.var == start.date & var_metric == "State space"),aes(x=var.value, y = metric.val,col="Start date"))+
  geom_point(data = filter(best_gam, thresh.var == last.date & var_metric == "State space"),aes(x=var.value, y = metric.val,col="End date"))+
  geom_point(data = filter(best_gam,transition == "trans" & var_metric == "State space"),aes(x=var.value, y = metric.val,col="Transition\ndates"))+
  ggrepel::geom_text_repel(data = filter(best_gam,transition == "trans" & var_metric == "State space"), aes(x=var.value, y = metric.val,label=thresh.var),force =1,nudge_x=0.55,nudge_y=-0.35,segment.linetype=2,min.segment.length = 0.1)+
  scale_colour_manual(values = c("blue","red","#FFE823"),breaks = c("Start date", "End date", "Transition\ndates"), name = NULL, guide = guide_legend(override.aes = list(size = 5))) +
  theme_classic() +
  theme(legend.text=element_text(size=14),
        strip.text.y = element_blank(),
        axis.title.y=element_blank(),
        strip.text.x = element_text(size=14),
        axis.title.x= element_text(size=14),
        legend.position = "top",
        panel.border = element_rect(linewidth = 1,fill = "transparent"))

##South

ts_plot + ss_plot

```


### Survival
1. recruits/spawner = s(year)
2. recruits/spawner = s(year) if BT < th


```{r rec_spwn_gam_model}
##North
n_dat <- ssb_recruit_bt_dat %>%
  filter(Region == "North") %>%
  mutate(rec_spawner = recruits/ssb) %>% 
  as.data.frame()
# n_1 <- mgcv::gam(log(recruits) ~ s(bt) + s(ssb) + s(year), data = n_dat)
# gratia::draw(n_1)
# gratia::appraise(n_1)
# summary(n_1)
# acf(resid(n_1))

# source("https://raw.githubusercontent.com/duncanobrien/ews-assessments/main/Code/threshold_gam.R")
# https://github.com/duncanobrien/ews-assessments/blob/main/Code/threshold_gam.R
  # data, #data.frame
  #           threshold.name, # colname of threshold variable to test within data
  #           nthd, #number of steps between lower & upper bound
  #           a = 0.2, #probability for lower bound quantile, 0:1
  #           b = 0.8, #probability for upper bound quantile, 0:1

# n_2 <- threshold.gam(log(recruits) ~ s(ssb) + s(year), nthd = 10, threshold.name = "bt", data = n_dat)
# summary(n_2)
# data = n_dat
cont_formula_ts <- as.formula("rec_spawner ~ s(year, bs='tp', k=3)")
thresh_formula_ts <- as.formula("rec_spawner ~ 1")

dens_gam_ts <- compare_gam(data = n_dat, 
                        cont_formula = formula(cont_formula_ts, method = "REML"),
                        thresh_formula =  formula(thresh_formula_ts, method = "REML"),
                        thresh.var = "year",
                        thresh.range = c(0.15,0.85), by = 1, k=3) #quartile range for threshold within 15-85%. Prevents bias towards thresholds at ends of time series


best_gam_ts <- predict_best_gam(object = dens_gam_ts) %>%
  dplyr::rename("metric.val" = rec_spawner) %>% 
  dplyr::mutate(threshold = tidyr::replace_na(as.character(threshold),"pre"),
                                                    start.date = dplyr::first(thresh.var),
                                                    last.date = dplyr::last(thresh.var)) %>% 
  pivot_longer(year, names_to = "var_metric", values_to = "var.value")


cont_formula_ss <- as.formula("rec_spawner ~ s(bt, bs='tp', k=3)")
thresh_formula_ss <- as.formula("rec_spawner ~ 1")
dens_gam_ss <- compare_gam(data = n_dat,
                        cont_formula = formula(cont_formula_ss, method = "REML"),
                        thresh_formula =  formula(thresh_formula_ss, method = "REML"),
                        thresh.var = "year", expl.var = "bt",
                        thresh.range = c(0.2,0.75), by = 1, k=3)

best_gam_ss <- predict_best_gam(object = dens_gam_ss) %>%
  dplyr::rename("metric.val" = rec_spawner) %>% 
  dplyr::mutate(threshold = tidyr::replace_na(as.character(threshold),"pre"),
                start.date = dplyr::first(thresh.var),
                last.date = dplyr::last(thresh.var)) %>% 
    pivot_longer(bt, names_to = "var_metric", values_to = "var.value")


best_gam <- bind_rows(best_gam_ts, best_gam_ss) %>% 
   mutate(var_metric = factor(ifelse(var_metric == "bt","Bottom Temperature","Year"),levels = c("Year","Bottom Temperature"), labels = c("Time series","State space")))

ts_plot <- ggplot(data = subset(best_gam, var_metric == "Time series"), aes(x=var.value,y=metric.val)) + 
  geom_point(aes(x=thresh.var, y = metric.val))+
  geom_path(aes(x=thresh.var, y = metric.val)) +  
  xlab("Year") + ylab("Scaled metric score") +
  facet_wrap(~ var_metric, scales = "free_x") +
  geom_line(data = filter(best_gam, threshold=="pre" & var_metric == "Time series"),aes(x=var.value,y=fit), col="blue",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="pre"& var_metric == "Time series"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#A1B4FE", col="#A1B4FE",alpha = 0.2) +
  geom_line(data = filter(best_gam, threshold=="post"& var_metric == "Time series"),aes(x=var.value,y=fit), col="red",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="post"& var_metric == "Time series"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#FFA6B9", col="#FFA6B9",alpha = 0.2)+
  geom_point(data = filter(best_gam, thresh.var == start.date & var_metric == "Time series"),aes(x=var.value, y = metric.val,col="Start date"))+
  geom_point(data = filter(best_gam, thresh.var == last.date & var_metric == "Time series"),aes(x=var.value, y = metric.val,col="End date"))+
  geom_point(data = filter(best_gam,transition == "trans" & var_metric == "Time series"),aes(x=var.value, y = metric.val,col="Transition\ndates"))+
  ggrepel::geom_text_repel(data = filter(best_gam, transition == "trans" & var_metric == "Time series"), aes(x=var.value, y = metric.val,label=thresh.var),force =1.5,nudge_x=5,nudge_y=-0.75,segment.linetype=2,min.segment.length = 0.1)+
  scale_colour_manual(values = c("blue","red","#FFE823"),breaks = c("Start date", "End date", "Transition\ndates"), name = NULL, guide = guide_legend(override.aes = list(size = 5))) +
  theme_classic() +
  scale_x_continuous(breaks = seq(1970,2010,20)) +
  theme(legend.text=element_text(size=14),
        strip.text.y = element_blank(),
        strip.text.x = element_text(size=14),
        axis.title = element_text(size=14),
        legend.position = "top",
        panel.border = element_rect(linewidth = 1,fill = "transparent"))
                                        

ss_plot <- ggplot(data = subset(best_gam,var_metric == "State space"), aes(x=var.value,y=metric.val)) + 
  geom_point(aes(x=var.value, y = metric.val))+
  geom_path(aes(x=var.value, y = metric.val)) +  
  xlab("Environment score") + ylab("Scaled metric score") +
  facet_wrap(~ var_metric, scales = "free_x") +
  geom_line(data = filter(best_gam, threshold=="pre" & var_metric == "State space"),aes(x=var.value,y=fit), col="blue",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="pre"& var_metric == "State space"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#A1B4FE", col="#A1B4FE",alpha = 0.2)+
  geom_line(data = filter(best_gam, threshold=="post"& var_metric == "State space"),aes(x=var.value,y=fit), col="red",linewidth=0.8, linetype = "solid")+
  geom_ribbon(data = filter(best_gam, threshold=="post"& var_metric == "State space"),aes(ymin = fit - (1.96 * ci),ymax = fit + (1.96 * ci)  ), fill = "#FFA6B9", col="#FFA6B9",alpha = 0.2)+
  geom_point(data = filter(best_gam, thresh.var == start.date & var_metric == "State space"),aes(x=var.value, y = metric.val,col="Start date"))+
  geom_point(data = filter(best_gam, thresh.var == last.date & var_metric == "State space"),aes(x=var.value, y = metric.val,col="End date"))+
  geom_point(data = filter(best_gam,transition == "trans" & var_metric == "State space"),aes(x=var.value, y = metric.val,col="Transition\ndates"))+
  ggrepel::geom_text_repel(data = filter(best_gam,transition == "trans" & var_metric == "State space"), aes(x=var.value, y = metric.val,label=thresh.var),force =1,nudge_x=0.55,nudge_y=-0.35,segment.linetype=2,min.segment.length = 0.1)+
  scale_colour_manual(values = c("blue","red","#FFE823"),breaks = c("Start date", "End date", "Transition\ndates"), name = NULL, guide = guide_legend(override.aes = list(size = 5))) +
  theme_classic() +
  theme(legend.text=element_text(size=14),
        strip.text.y = element_blank(),
        axis.title.y=element_blank(),
        strip.text.x = element_text(size=14),
        axis.title.x= element_text(size=14),
        legend.position = "top",
        panel.border = element_rect(linewidth = 1,fill = "transparent"))

##South

ts_plot + ss_plot

```

