---
title: "Run Commercial + Recreational Indicator Functions"
author: "Stephanie Owen"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
`%>%` <- magrittr::`%>%`
```

## Recreational Trips = 'create_rec_trips'

Pull in directed trip files and run rec trips function
```{r cars}
path = (here::here('data-raw/mrip_directed_trips'))

files <- list.files(path = path, pattern = glob2rx('mrip*.csv'),
                       full.names = TRUE)

rec_trips <- create_rec_trips(files = files)
```

Filter to N stock only and sum
```{r}
rec_trips2 <- rec_trips %>%
  dplyr::filter(STATE %in% c('MAINE',  'CONNECTICUT',
                                 'MASSACHUSETTS',
                                 'NEW HAMPSHIRE',
                                 'NEW JERSEY',
                                 'NEW YORK',
                                 'RHODE ISLAND',
                                 'MARYLAND',
                                 'DELAWARE',
                                 'NORTH CAROLINA')) %>% # northern stock only (Northern MAB))
  dplyr::group_by(YEAR) %>%
  dplyr::summarise(DATA_VALUE = sum(as.numeric(DATA_VALUE), na.rm = TRUE)) %>%
  dplyr::mutate(CATEGORY = "Recreational",
         INDICATOR_TYPE = "Socioeconomic",
         INDICATOR_NAME = "total_recreational_trips_n")
```

## Recreational Effort = 'create_prop_sp_trips'

Pull in effort file and run effort function - this calculatess all recreational effort to get the proportion of recreational effort for BSB only
```{r}
mrip_effort <- read.csv(here::here('data-raw/mrip_effort_series.csv'),
                        skip = 44, # was 24 is now 44
                        na.strings = ".")

prop_sp_trips <- create_prop_sp_trips(total = mrip_effort, sp = rec_trips)
```

## Total Recreational Catch = 'create_total_rec_catch'

Read in MRIP catch data and run total rec catch function
```{r}
mrip_catch <- read.csv(here::here('data-raw/mrip_BLACK_SEA_BASS_catch_series.csv'),
                                  skip = 46, # of rows you want to ignore
                                  na.strings = ".")

total_rec_catch <- create_total_rec_catch(mrip_catch)
```

## Total Recreational Landings = 'create_total_rec_landings'

Read in MRIP harvest data and run total rec landings function
```{r}
mrip_landing <- read.csv(here::here('data-raw/mrip_BLACK_SEA_BASS_harvest.csv'),
                         skip = 46, # of rows you want to ignore
                         na.strings = ".")

total_rec_landings <- create_total_rec_landings(mrip_landing)
```

## Format Commercial Data for Plotting

```{r}
dat <- read.csv(here::here('data-raw/SOCIEOECONOMIC_COMMERCIAL_INDICATORS_BLACK_SEABASS_FINAL.csv'))

landings <- dat[1:28, ] %>% #filter landings only rows
  dplyr::mutate(DATA_VALUE = (DATA_VALUE)/1000000) #divide to reduce # of zeros on y axis label

vessels <- dat[29:56, ] #filter vessels only rows
```

