---
title: "Black Sea Bass (\\protect\\textit{Centropristis striata}) \\linebreak Ecosystem & Socioeconomic Profile Report Card"
format: 
  pdf:
    margin-left: "0.5in"
    margin-right: "0.5in"
    margin-top: "0.5in"
    margin-bottom: "0.5in"
    fontsize: "10pt"
    linestretch: 1
    indent: true
    include-in-header: tex_template.tex    
    keep-tex: true
    papersize: letter
params:
  widths: !expr c(0.9, 0.75, 2.5, 3)
  img1: !expr here::here("images/bsb spatiotemporal conceptual model.png")
  img2: !expr here::here("images/new_condition.png")
  img3: !expr here::here("images/risk_plot.png")
  tbl_file: !expr here::here("docs/bsb_indicator_table.csv")
  img_dir: !expr here::here("images")
latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
devtools::load_all(here::here("../READ-EDAB-NEesp2"))
# here::i_am("inst/esp_template/06_docs/report card pg 1.qmd")
```

\backgroundsetup{
  scale=1,
  angle=0,
  opacity=0,
  contents={\includegraphics[width=\paperwidth,height=\paperheight]{bg_pg1.jpg}}
 }
\BgThispage

\vspace{-2.5cm}
\section{Spring 2025}

This is a short-form update to the full Ecosystem and Socioeconomic Profile (Tabandera et al., 2024) highlighting the recent status of environmental, ecological, and socioeconomic factors. Black sea bass is an important Mid-Atlantic stock with high commercial value and recreational engagement. Overfishing is not occurring and the stock is not overfished. Winter bottom temperature is used in the stock assessment model as a factor that influences recruitment  to incorporate the observed link between cold temperature and smaller year classes (NEFSC 2023).

<!-- \vspace{-0.5cm} -->

::: {layout="[ [0.4, 0.025, 0.575], [1], [0.575, 0.025, 0.4]] "}

:::: {}
![](`r params$img1`)
::::

:::: {}
\hfill
::::

:::: {}
\section{2024 in Review}

\subsection{Fishing Community Observations (MAFMC 2024a-c)}

-   Steady or increasing availability
-   Restrictive and complex regulations limit fishing opportunities
-   Concerns about high discards

\vspace{-0.25cm}
\subsection{Commercial Fishery}

-   Number of active vessels declined in 2024, but total landed pounds increased from 2023 
-   Total revenue decreased slightly along with average prices (\$/lb)
-   Average revenue per vessel increased

\vspace{-0.25cm}
\subsection{Recreational Fishery}

-   Number of targeted trips, catch, and landings all down from 2023 (MAFMC 2024a)
-   However, number of trips is still above the historic average
-   Recreational catch-per-angler index not yet updated for 2024

\vspace{-0.25cm}
\subsection{Ecosystem}

-   The stock assessment models the stock as two subunits, divided at the Hudson Canyon (see right)
-   Cold winter in the north but near average in the south
-   Poor or below average fish condition (i.e., weight at a given length) in recent years
::::

:::: {}
\vspace{0.25cm}
\section{Key Points from the Mid-Atlantic Risk Assessment}
\vspace{0.05cm}
<!-- \vspace{0.15cm} -->
::::

:::: {}
\raggedright

<!-- maybe this is being centered because it's the only section without a header? -->

According to the [Mid-Atlantic 2024 EAFM risk assessment update](https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/6747560a3cf66936045e5547/1732728332670/05_EAFM+Risk+Assessment.pdf) (MAFMC 2024d), Black Sea Bass scored high and/or moderately high risk in the following elements (elements with low or low-moderate scores not described here):

-   Moderate-high to high risk to the stock due to:
    -   Very high exposure to changes in climate
    -   Observed and potential changes in distribution; northward shift into the Gulf of Maine
    -   Dependence on threatened estuarine habitat
    -   Decline in the biomass of benthic invertebrate prey
    -   Decline in black sea bass body condition
-   High risk to the recreational fishery due to:
    -   Catch exceeding harvest limits in several years
    -   High regulatory complexity; frequent changes and varying interstate regulations; regulatory changes in allocations
-   Moderate-high risk to the commercial fishery due to:
    -   Commercial revenue in wind development areas
    -   High discards & discard mortality


```{r, eval = FALSE}
devtools::load_all("../READ-EDAB-NEesp2")
risk_fig <- NEesp2::plot_risk(risk_elements = tibble::tibble(stock = c(4, 0, 2, 2),
                                                              recreational = c(1, 1, 1, 3),
                                                              commercial = c(1, 4, 2, 0)))
ggplot2::ggsave(risk_fig, 
                filename = here::here("images/risk_plot.png"),
                width = 4,
                height = 2.5,
                dpi = 300)
```
::::

:::: {}
\hfill
::::

:::: {}
![](`r params$img2`) 
![](`r params$img3`)
::::

:::

\newpage

\backgroundsetup{
  scale=1,
  angle=0,
  opacity=0,
  contents={\includegraphics[width=\paperwidth,height=\paperheight]{bg_pg2.jpg}}
 }
\BgThispage

<!-- not sure why this background isn't working -->

<!-- \newgeometry{top=0.75in, left=0.75in, right=0.75in, bottom=0.75in} -->

```{r}
data <- NEesp2::format_tbl_data(file = params$tbl_file,
                            term_year = params$terminal_year,
                            dir = params$img_dir)
```

```{r}
tbl <- NEesp2::rpt_card_table(data,
                       widths = params$widths)
tbl <- flextable::set_header_labels(tbl, `Time Series` = "Time Series*")

tbl <- flextable::font(tbl, fontname = "Times New Roman", part = "body")
tbl <- flextable::font(tbl, fontname = "Times New Roman", part = "header")
```

```{r}
# this doesn't seem to do anything
tbl <- flextable::height_all(tbl, height = 2, part = "body", unit = "in")
tbl <- flextable::hrule(tbl, rule = "atleast", part = "body")

tbl

# flextable::save_as_image(tbl, here::here("images/indicator_table.png"))
```
\vspace{-0.4cm}
\footnotesize \* The y-axis units are included in the "Indicator" column of the table. In all figures, the dashed line represents the time series mean, and the solid green lines indicate ± 1 standard deviation. Commercial data were derived from the commercial dealer database hosted at the Greater Atlantic Regional Office. All dollar values have been adjusted to 2024 real dollars using the [Gross Domestic Implicit Price Deflator](https://fred.stlouisfed.org/series/GDPDEF).\newline\newline

\centering\normalsize

We welcome your observations! Please contact northeast.ecosystem.highlights@noaa.gov with any on-the-water insights or changes observed in the black sea bass fishery and nefsc.esp.leads@noaa.gov with questions or comments on the information presented in this report.

\newpage
\newgeometry{top=1in, left=1in, right=1in, bottom=1in}

\section{References}

\raggedright

Fratantoni, P.S., Holzworth-Davis, T., & Taylor, M.H. (2015). Description of oceanographic conditions on the Northeast US Continental Shelf during 2014. US Dept Commer, Northeast Fisheries Science Center. Ref Doc. 15-21; 41 p. http://dx.doi.org/10.7289/V5SQ8XD2 \newline\newline

Jean-Michel, L., Eric, G., Romain, B.-B., Gilles, G., Angélique, M., Marie, D., Clément, B., Mathieu, H., Olivier, L. G., Charly, R., Tony, C., Charles-Emmanuel, T., Florent, G., Giovanni, R., Mounir, B., Yann, D., & Pierre-Yves, L. T. (2021). The Copernicus Global 1/12° Oceanic and Sea Ice GLORYS12 Reanalysis. Frontiers in Earth Science, 9, 698876. https://doi.org/10.3389/feart.2021.698876 \newline\newline

MAFMC. (2024a). Black Sea Bass Fishery Information Document. https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/6683032fa306b9070227348b/1719862063436/2024_BSB-fishery-info-doc.pdf \newline\newline

MAFMC. (2024b). Summer Flounder, Scup, and Black Sea Bass Fishery Performance Report. https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/6697d5dde4c39b20168045b7/1721226717179/b_SFSBSB-FPR-Jul-2024.pdf \newline\newline

MAFMC. (2024c). Summer Flounder, Scup, and Black Sea Bass Monitoring Committee (MC) November 19, 2024 Webinar Meeting Summary 2025 Recreational Management Measures. https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/674755f846e791454569703f/1732728312681/01_2025+SF+Rec+Measures.pdf \newline\newline

MAFMC. (2024d). Mid-Atlantic EAFM Risk Assessment: 2024. https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/6747560a3cf66936045e5547/1732728332670/05_EAFM+Risk+Assessment.pdf \newline\newline

NEFSC. (2023). Report of the Black Sea Bass (Centropristis striata) Research Track Stock Assessment Working Group. https://d23h0vhsm26o6d.cloudfront.net/11a.-2023_BSB_UNIT_RTWG_Report_V2_12_2_2023.pdf \newline\newline

Tabandera, R., Tyrell, A., McMahan, M., Perez, P., Hankowsky, K., & Large, S. (2024). Black sea bass ecosystem considerations and indicator development.  https://doi.org/10.25923/ez9g-af05


