---
title: "Black sea bass shapefiles"
author: "Abigail Tyrell"
date: "11/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
`%>%` <- magrittr::`%>%`
```

The purpose of this document is to create shapefiles for the southern MAB, northern MAB, and GB/GOM regions of black sea bass habitat.

First, here are the strata:
```{r}
strata <- readxl::read_excel(here::here("black-sea-bass/BSB.2021RT.NEFSC.Strata.Specifications.By.Region.xlsx")) %>%
  # dplyr::select(-Series) %>%
  dplyr::distinct()
```

Use these strata with the NEFSC bottom trawl strata shapefile to create a black sea bass shapefile. Split the northern subgroup into NMAB and GB/GOM. Not sure why this can't run in Rmarkdown, so saving output and reading in.
```{r, eval = FALSE}
bsb_shape <- NEesp::shape %>%
  dplyr::select(STRATA, geometry) %>%
  dplyr::filter(STRATA %in% strata$Stratum) %>%
  dplyr::full_join(strata,
                   by = c("STRATA" = "Stratum")) %>%
  dplyr::mutate(Region = ifelse(Region == "North" &
                       (STRATA < 1090 |
                         (STRATA > 3000 &
                            STRATA < 3460)), "NMAB", Region)
                )
saveRDS(bsb_shape, here::here("black-sea-bass/bsb_shape.RDS"))
```

```{r}
bsb_shape <- readRDS(here::here("black-sea-bass/bsb_shape.RDS"))
```

Plot to make sure everything looks right:

```{r, echo = FALSE}
xlims <- c(-81, -66)
ylims <- c(34, 45)
crs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

ne_countries <- rnaturalearth::ne_countries(
  scale = 10,
  continent = "North America",
  returnclass = "sf"
) %>%
  sf::st_transform()

ne_states <- rnaturalearth::ne_states(
  country = "united states of america",
  returnclass = "sf"
) %>% sf::st_transform()

# plot
p1 <- ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = bsb_shape,
    ggplot2::aes(fill = Region)
  ) +
  ggplot2::geom_sf(
    data = ne_countries,
    color = "grey60",
    size = 0.25
  ) +
  ggplot2::geom_sf(
    data = ne_states,
    color = "grey60",
    size = 0.05
  ) +
  ggplot2::coord_sf(
    crs = crs,
    xlim = xlims,
    ylim = ylims
  ) +
  ggthemes::theme_map() +
  ggplot2::theme(legend.direction = "horizontal")
print(p1)
```

Simplify the shapefile:
```{r}
bsb_shape2 <- bsb_shape %>%
  dplyr::group_by(Region) %>%
dplyr::summarise(geometry = sf::st_union(geometry))

bsb_shape2
```

Plot the simplified shapefile:

```{r, echo = FALSE}
p1 <- ggplot2::ggplot() +
  ggplot2::geom_sf(
    data = bsb_shape2,
    ggplot2::aes(fill = Region)
  ) +
  ggplot2::geom_sf(
    data = ne_countries,
    color = "grey60",
    size = 0.25
  ) +
  ggplot2::geom_sf(
    data = ne_states,
    color = "grey60",
    size = 0.05
  ) +
  ggplot2::coord_sf(
    crs = crs,
    xlim = xlims,
    ylim = ylims
  ) +
  ggthemes::theme_map() +
  ggplot2::theme(legend.direction = "horizontal")
print(p1)
```

Save both shapefiles:
```{r}
sf::write_sf(bsb_shape, here::here("black-sea-bass/full_strata_bsb.shp"))
sf::write_sf(bsb_shape2, here::here("black-sea-bass/simplified_bsb.shp"))
```
