---
title: "Black Sea Bass (\\protect\\textit{Centropristis striata}) \\linebreak Snapshot Ecosystem & Socioeconomic Profile"
format: 
  pdf:
    margin-left: "0.5in"
    margin-right: "0.5in"
    margin-top: "0.5in"
    margin-bottom: "0.5in"
    fontsize: "10pt"
    linestretch: 1
    indent: true
    include-in-header: tex_template.tex    
    keep-tex: true
    papersize: letter
params:
  widths: !expr c(0.9, 0.75, 2, 3)
  img1: !expr here::here("images/bsb spatiotemporal conceptual model.png")
  img2: !expr here::here("images/new_condition.png")
  img3: !expr here::here("images/risk_plot.png")
  tbl_file: !expr here::here("docs/bsb_indicator_table.csv")
  img_dir: !expr here::here("images")
  terminal_year: 2024
latex_engine: xelatex
bibliography: bsb_references.bib
nocite: |
  @jeanmichel_copernicus_2021, @fratantoni_description_2015
csl: cambridge-university-press-numeric.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
devtools::load_all(here::here("../READ-EDAB-NEesp2"))
# here::i_am("inst/esp_template/06_docs/report card pg 1.qmd")
```

\backgroundsetup{
  scale=1,
  angle=0,
  opacity=0,
  contents={\includegraphics[width=\paperwidth,height=\paperheight]{bg_pg1.jpg}}
 }
\BgThispage

\vspace{-2.5cm}
\section{Spring 2025}

This is a short-form update to the full Ecosystem and Socioeconomic Profile [@tabandera_black_2024] highlighting the recent status of environmental, ecological, and socioeconomic factors. Black sea bass is an important Mid-Atlantic stock with high commercial value and recreational engagement. Overfishing is not occurring and the stock is not overfished. Winter bottom temperature is used in the stock assessment model as a factor that influences recruitment  to incorporate the observed link between cold temperature and smaller year classes [@nefsc_report_2023].

\vspace{-0.25cm}

::: {layout="[ [0.4, 0.025, 0.575], [1], [0.575, 0.025, 0.4]] "}

:::: {}
![](`r params$img1`)
::::

:::: {}
\hfill
::::

:::: {}
\section{2024 in Review}
\vspace{-0.25cm}

\subsection{Fishing Community Observations}

-   Steady or increasing availability 
-   Concerns about high discards
-   Restrictive and complex regulations limit fishing opportunities
-   For additional information, see @mafmc_black_2024, @mafmc_summer_2024, and @mafmc_summer_2024b

\vspace{-0.25cm}
\subsection{Commercial Fishery}

-   Number of active vessels declined in 2024, but total landed pounds increased from 2023 
-   Total revenue decreased slightly along with average prices (\$/lb)
-   Average revenue per vessel increased

\vspace{-0.25cm}
\subsection{Recreational Fishery}

-   Targeted trips, catch, and landings all down from 2023 [@mafmc_black_2024]
-   However, number of trips is still above the historic average
-   Recreational catch-per-angler index not yet updated for 2024

\vspace{-0.25cm}
\subsection{Ecosystem}

-   The stock assessment models the stock as two subunits, divided at the Hudson Canyon (see right)
-   Cold winter in the north but near average in the south
-   Poor or below average fish condition (i.e., weight at a given length; see below)
::::

:::: {}
\vspace{0.25cm}
\section{Key Points from the Mid-Atlantic Risk Assessment}
\vspace{0.05cm}
<!-- \vspace{0.15cm} -->
::::

:::: {}
\raggedright

<!-- maybe this is being centered because it's the only section without a header? -->
\subsection{Moderate-high to high risk of the stock not achieving optimal biomass due to:}

-   Very high exposure to changes in climate
-   Observed and potential changes in distribution; northward shift into the Gulf of Maine
-   Dependence on threatened estuarine habitat
-   Decline in the biomass of benthic invertebrate prey
-   Decline in black sea bass body condition (see right)

\vspace{-0.25cm}
\subsection{High risk of the recreational fishery not achieving optimal yield due to:}

-   Catch exceeding harvest limits in several years
-   High regulatory complexity; frequent changes and varying interstate regulations; regulatory changes in allocations

\vspace{-0.25cm}
\subsection{Moderate-high risk of the commercial fishery not achieving optimal yield due to:}

-   Commercial revenue in wind development areas
-   High discards & discard mortality

\vspace{-0.25cm}
Please see the [Mid-Atlantic 2024 EAFM risk assessment update](https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/6747560a3cf66936045e5547/1732728332670/05_EAFM+Risk+Assessment.pdf) for more details and explanation of the low and low-moderate risks to black sea bass and its associated fisheries.


```{r, eval = FALSE}
devtools::load_all("../READ-EDAB-NEesp2")
risk_fig <- NEesp2::plot_risk(risk_elements = tibble::tibble(stock = c(4, 0, 2, 2),
                                                              recreational = c(1, 1, 1, 3),
                                                              commercial = c(1, 4, 2, 0)))
ggplot2::ggsave(risk_fig, 
                filename = here::here("images/risk_plot.png"),
                width = 4,
                height = 2.5,
                dpi = 300)
```
::::

:::: {}
\hfill
::::

:::: {}
![](`r params$img2`) 
![](`r params$img3`)
::::

:::

\newpage

\backgroundsetup{
  scale=1,
  angle=0,
  opacity=0,
  contents={\includegraphics[width=\paperwidth,height=\paperheight]{bg_pg2.jpg}}
 }
\BgThispage

<!-- not sure why this background isn't working -->

\newgeometry{top=0.75in, left=0.75in, right=0.75in, bottom=0.75in}

```{r tbl1, include = FALSE}
data <- NEesp2::format_tbl_data(file = params$tbl_file,
                            term_year = params$terminal_year,
                            dir = params$img_dir)
```

```{r tbl2, include = FALSE}
tbl <- NEesp2::rpt_card_table(data,
                       widths = params$widths)
tbl <- flextable::set_header_labels(tbl, `Time Series` = "Time Series*")

tbl <- flextable::font(tbl, fontname = "Times New Roman", part = "body")
tbl <- flextable::font(tbl, fontname = "Times New Roman", part = "header")
```

```{r tbl3, include = FALSE}
# this doesn't seem to do anything
# tbl <- flextable::height_all(tbl, height = 2, part = "body", unit = "in")
# tbl <- flextable::hrule(tbl, rule = "atleast", part = "body")

# add superscripts
tbl <- flextable::compose(tbl, 
                          part = "body", 
                          i = 1, 
                          j = 1, 
                          value = flextable::as_paragraph("Mean winter (Feb-Mar) bottom temperature (\u00B0C)"))
tbl <- flextable::compose(tbl, 
                          part = "body", 
                          i = 2, 
                          j = 1, 
                          value = flextable::as_paragraph("Shelf water volume (km", 
                                                          flextable::as_sup("3"), 
                                                          ")"))

# tbl <- ftExtra::colformat_md(tbl)
tbl
# flextable::save_as_image(tbl, here::here("images/indicator_table.png"))
```

```{r}
tbl
```

\vspace{-0.4cm}
\footnotesize \* The y-axis units are included in the "Indicator" column of the table. In all figures, the dashed line represents the time series mean, and the solid green lines indicate ± 1 standard deviation. Commercial data were derived from the commercial dealer database hosted at the Greater Atlantic Regional Office. All dollar values have been adjusted to 2024 real dollars using the [Gross Domestic Implicit Price Deflator](https://fred.stlouisfed.org/series/GDPDEF).\newline

\centering\normalsize

We welcome your observations! Please contact [northeast.ecosystem.highlights@noaa.gov](northeast.ecosystem.highlights@noaa.gov) with any on-the-water insights or changes observed in the black sea bass fishery and [nefsc.esp.leads@noaa.gov](nefsc.esp.leads@noaa.gov) with questions or comments on the information presented in this report.

\newpage
\newgeometry{top=1in, left=1in, right=1in, bottom=1in}

# References
::: {#refs}
:::

<!-- Fix ref ordering -->
fix ref order (delete me): [@jeanmichel_copernicus_2021; @fratantoni_description_2015]
